---
title: "Data Manipulation"
author: "ak5357"
date: "2024-09-19"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data.

Import the two datasets that we're going to manipulate.

```{r}
litters_df = read_csv("data/FAS_litters.csv", na = c("NA", "", "."))
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("data/FAS_pups.csv", na = c("NA", "", ".", " "))
pups_df = janitor::clean_names(pups_df)

```

## `select`
Use `select()` to select columns, or variables. Choose a dataframe to select from and list the columns you want to see, it will show them to you.

Certain columns:
```{r}
select(litters_df, group, litter_number, gd0_weight)
```


Range of columns:
```{r}
select(litters_df, group:gd18_weight)
```


Exclude columns:
```{r}
select(litters_df, -pups_survive)
```


Exclude range of columns:
```{r}
select(litters_df, -(group:gd18_weight))
```


Conditionally select based on starting chars:
```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))
```


Rename variables as you go:
```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)
```


Reorder columns:
```{r}
select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight) #assumes everything else comes after this
```


Learning Assessment:
```{r}
select(pups_df, litter_number, sex, pd_ears)
```



## `filter`
Use this to filter out rows conditionally based on some variables. It's case sensitive.

Equal to:
```{r}
filter(litters_df, gd_of_birth == 20)
filter(litters_df, gd_of_birth == 19)
```


Greater than or equal to:
```{r}
filter(litters_df, pups_born_alive > 8)
filter(litters_df, pups_born_alive >= 8)
```


Not equal to:
```{r}
filter(litters_df, pups_born_alive != 9)
```


In:
```{r}
filter(litters_df, group == "Con7")
filter(litters_df, group %in% c("Low7", "Low8"))
```


Multiple conditions (AND):
```{r}
filter(litters_df, group %in% c("Low7", "Low8"), pups_born_alive == 8)
```


Multiple conditions (OR):
```{r}
filter(litters_df, group %in% c("Low7", "Low8") | pups_born_alive == 8)
```


Get rid of missing data:
```{r}
drop_na(litters_df)
```

Learning Assessment:
```{r}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk < 11, sex == 2)
```



## `mutate`

One variable at a time:
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))
```


Multiple variables at once:
```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```



## `arrange`

Reorder the rows:

```{r}
arrange(litters_df, gd0_weight) #ascending

arrange(litters_df, desc(gd0_weight)) #descending

arrange(litters_df, pups_born_alive, gd0_weight)
```


## PIPING

Avoid repeat-overwriting or nesting functions. Use piping instead!

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "", ".", " ")) |> 
  janitor::clean_names() |> 
  select(-pups_born_alive) |>
  mutate(
    wt_gain = gd18_weight - gd0_weight
  )
```


```{r}
read_csv("data/FAS_litters.csv", na = c("NA", "", ".", " ")) |> 
  janitor::clean_names() |> 
  mutate(wt_gain = gd18_weight - gd0_weight) |> 
  lm(wt_gain ~ pups_born_alive, data = _)

```



## Data Export
```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c("NA", "", ".", " ")) |> 
  janitor::clean_names() |> 
  mutate(wt_gain = gd18_weight - gd0_weight) |> 
  select(-pups_born_alive) |> 
  filter(group == "Con7")

write_csv(litters_df, "data/cleaned_fas_litters.csv")
```
















