---
title: "Data Import"
author: "ak5357"
date: "2024-09-17"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data.


## Import the FAS Litters CSV

Import the dataset and clean the names.

```{r}
litters_df = read_csv(
  file = "data/FAS_litters.csv",
  na = c("NA", "", "."),
  col_types = cols(
    Group = col_factor()
  )
  )
litters_df = janitor::clean_names(litters_df)
```

## Look at the dataset

```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)
```


## Import the FAS Pups CSV

Use relative path, much easier to use and to share with collaborators.

```{r}
pups_df = read.csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)

head(pups_df)
```


## Import MLB 2011 Summary Data

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


## Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
head(pulse_df)
```


## Never use read.csv()

read_csv is a Tibble function, whereas read.csv is local to R. The tibble version is faster, more memory efficient, and has some helpful abilities and guardrails.

```{r}
litters_df = read.csv("data/FAS_litters.csv")
litters_df$L
```

























