---
title: "Reading Data from the Web"
author: "ak5357"
date: "2024-10-10"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
```


### Load data from web

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

drug_use_html
```


### Extracting data from HTML

To get a certain table from an HTML, you can use `first()` to get the first table, `last()` to get the last one, and `nth(x, n)` to get others.

```{r}
# Imports all tables from the HTML
drug_use_html |>
  html_table()

# Import only the first table
table_marj = 
  drug_use_html |> 
  html_table() |> 
  first() |> 
  slice(-1) #removes first column from table

table_marj
```


### Learning Assessment: NYC Cost of Living

Create a data frame that contains the cost of living table for New York from this page: https://www.bestplaces.net/cost_of_living/city/new_york/new_york.

```{r}
nyc_url = "https://www.bestplaces.net/cost_of_living/city/new_york/new_york"
nyc_cost_html = read_html(nyc_url)

nyc_cost_df =
  nyc_cost_html |> 
  html_table(header = TRUE) |> 
  first()

nyc_cost_df
```

### CSS Selectors

**Example: Star Wars Movies on IMDB**
```{r}
# Star Wars Movies
swm_url = "https://www.imdb.com/list/ls070150896/"
swm_html = read_html(swm_url)

# Movie title
swm_title_vec = swm_html |> 
  html_elements(".dli-title") |> 
  #html_elements(".ipc-title-link-wrapper .ipc-title__text") |> 
  html_text()

# Runtime
swm_runtime_vec = swm_html |> 
  html_elements(".dli-title-metadata-item:nth-child(2)") |> 
  html_text()

# Parental Guidance
swm_pg_vec = swm_html |> 
  html_elements(".dli-title-metadata-item:nth-child(3)") |> 
  html_text()

# Score
swm_score_vec = swm_html |> 
  html_elements(".metacritic-score-box") |> 
  html_text()

# All movie data in a df
swm_df = 
  tibble(
    title = swm_title_vec,
    runtime = swm_runtime_vec,
    pg = swm_pg_vec,
    score = swm_score_vec
  )

swm_df

```


**Example: Book Titles**
```{r}
books_url = "https://books.toscrape.com/"
books_html = read_html(books_url)

books_titles = 
  books_html |> 
  html_elements("h3") |> 
  html_text2()

books_stars = 
  books_html |>
  html_elements(".star-rating") |>
  html_attr("class")

books_price = 
  books_html |>
  html_elements(".price_color") |>
  html_text()

books_df = tibble(
  title = books_titles,
  stars = books_stars,
  price = books_price
)

books_df
```

### Scrape API
**Example: NYC Open Data**
```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") |> 
  content()

nyc_water
```

**Example: Data.gov Behavioral Risk Factors dataset**
```{r}
brfss_df = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) |> #parameter for selecting more than the default given rows from this specific API
  content()

brfss_df
```

**Example: Pokemon API**
This one spits out data that is definitely not a CSV.
```{r}
pokemon =
  GET("https://pokeapi.co/api/v2/pokemon/ditto") |> 
  content()

pokemon$height
pokemon$abilities
```





