---
title: "Statistical Learning"
author: "ak5357"
date: "2024-11-21"
output: html_document
---

```{r include = FALSE}
library(tidyverse)
library(glmnet)
library(viridis)

set.seed(1)
```

Import and clean birthweight data.
```{r}
bwt_df =
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = 
      case_match(babysex,
          1 ~ "male",
          2 ~ "female"),
    frace = 
      case_match(frace,
          1 ~ "white",
          2 ~ "black",
          3 ~ "asian",
          4 ~ "puerto rican",
          8 ~ "other"),
    frace = fct_infreq(frace),
    mrace = 
      case_match(mrace,
          1 ~ "white",
          2 ~ "black",
          3 ~ "asian",
          4 ~ "puerto rican",
          8 ~ "other"),
    mrace = fct_infreq(mrace)) |> 
  sample_n(200) #Take sample of 200
```


Construct inputs for `glmnet`

* design matrix = all predictors that will go into model
* output vector

```{r}
x = model.matrix(bwt ~ ., data = bwt_df)[, -1]
y = bwt_df |> pull(bwt)

x
y
```

Fit lasso for several lambdas

* Jeff tried a bunch of lamda values and settled on this range
* glmnet is extremely computationally efficient
```{r}
lambda = 10^seq(-2, 2.75, by = 0.1)

# Get LASSO fit
lasso_fit = glmnet(x = x, y = y, lambda = lambda)
lasso_fit

# Perform cross validation
lasso_cv = cv.glmnet(x = x, y = y, lambda = lambda)
lasso_cv

# Get lambda for best fit
lambda_optimal = lasso_cv[["lambda.min"]]
lambda_optimal
```

Usual lasso plot
```{r}
lasso_fit |> 
  broom::tidy() |>
  filter(term != "(Intercept)") |> 
  select(term, lambda, estimate) |> 
  complete(term, lambda, fill = list(estimate = 0)) |>
  ggplot(aes(x = lambda, y = estimate, group = term, color = term)) +
  geom_line() +
  geom_vline(xintercept = lambda_optimal, color = "black")
```

Final model with best lambda.
```{r}
final_lasso_fit =
  glmnet(x = x, y = y, lambda = lambda_optimal)

final_lasso_fit |> 
  broom::tidy()
```

Look at CV results
```{r}
lasso_cv |> 
  broom::tidy() |> 
  ggplot(aes(x = log(lambda, base = 10), y = estimate)) +
  geom_point
```

How is this different from regression modeling?

* Exclusively focused on prediction accuracy.
* Not going to give same regression coefficients.
















