---
title: "Strings and Factors"
author: "ak5357"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(p8105.datasets)
library(stringr)
library(tidyverse)
library(dplyr)
library(forcats)
```


## **Strings and Regex**

**Searching for substring**

When doing string operations, need to have precise, exact matches when searching.
```{r}
string_vec = c("my", "name", "is", "arghya")

str_detect(string_vec, "arghya") #returns array of booleans

str_replace(string_vec, "arghya", "Arghya")

str_replace(string_vec, "e", "E")
```

**Specify location of search string within source strings**
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
)

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think") # ^ at beginning --> starts with

str_detect(string_vec, "i think$") # $ at end --> ends with
```

**Searching pattern allowing multiple characters**
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"pumpkin")

str_detect(string_vec,"Pumpkin")

str_detect(string_vec, "PUMPKIN")

str_detect(string_vec,"[Pp]umpkin") #allow either option
```

**Searching for multiple consecutive characters**
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z][a-zA-Z]") # starts with number then two letters
```

**Using . as wild card to capture all characters**
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # . is a special character that matches everything
```

**Escape character**
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[") # find [

str_detect(string_vec, "[\\[\\]]") # find either [ or ]
```

## **Factors ... **

```{r}
sex_vec = factor(c("male", "male", "female", "female"))

as.numeric(sex_vec)

```

**Relevel**
```{r}
sex_vec = fct_relevel(sex_vec, "male")
```

## Web Scraping example

**Get data**
```{r}
# Declare and get url
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
drug_use_html = read_html(url)

# Import only the first table
marj_use_df = 
  drug_use_html |> 
  html_table() |> 
  first() |> 
  slice(-1) #removes first column from table
```

**Tidy data**
```{r}
marj_use_df |> colnames()

marj_use_df = marj_use_df |> 
  select(-contains("P value")) |> 
  pivot_longer(
    cols = -State,
    names_to = "age_year",
    values_to = "percent"
  ) |> 
  separate(age_year, into = c("age", "year"), sep = "\\(") |> 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_remove(percent, "[a-c$]"), # we expect a-c, so we want there to be an error if we see any other characters that we don't expect
    percent = as.numeric(percent)
  )
```

**Plot**
```{r}
marj_use_df |> 
  filter(age == "12-17") |> 
  mutate(
    State = fct_reorder(State, percent)
  ) |> 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  labs(
    title = "Marijuana use by 12-17 year-olds (%)",
    subtitle = "by U.S. State",
    x = "State",
    y = "Percent (%)",
    color = "Year"
  ) +
  theme(
    
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
  #coord_flip()
```


## NYC Restaurant Inspections

```{r}
data("rest_inspec")

rest_inspec |> colnames()

rest_inspec |> 
  count(boro, grade) |> 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
```



```{r}
rest_inspec = rest_inspec |> 
  filter(
    str_detect(grade, "[A-C]"),
    !(boro == "Missing")
  )
```

**On reordering factors**
```{r}
rest_inspec |> 
  mutate(dba = str_to_sentence(dba)) |> 
  filter(str_detect(dba, "Pizza")) |> 
  mutate(
    # REORDER FACTORS
    boro = fct_infreq(boro), #change to a factor variable
    
    # REPLACE/RENAME STRINGS
    #boro = str_replace(boro, "MANHATTAN", "THE CITY") #str_replace changes it back to a string variable
    
    # REPLACE/RECODE FACTORS
    #boro = fct_recode(boro, "THE CITY" = "MANHATTAN")
  ) |> 
  #mutate( boro = fct_relevel(boro, "STATEN ISLAND")) |>
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

**Nonsense Regression Model Example**
```{r}
rest_inspec |> 
  mutate(dba = str_to_sentence(dba)) |> 
  filter(str_detect(dba, "Pizza")) |> 
  mutate(boro = fct_infreq(boro)) |> 
  lm(zipcode ~ boro, data = _)
```










