---
title: "Strings and Factors"
author: "ak5357"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
library(rvest)
library(p8105.datasets)
library(stringr)
library(dplyr)
library(forcats)
```


## **Strings and Regex**

**Searching for substring**

When doing string operations, need to have precise, exact matches when searching.
```{r}
string_vec = c("my", "name", "is", "arghya")

str_detect(string_vec, "arghya") #returns array of booleans

str_replace(string_vec, "arghya", "Arghya")

str_replace(string_vec, "e", "E")
```

**Specify location of search string within source strings**
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
)

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think") # ^ at beginning --> starts with

str_detect(string_vec, "i think$") # $ at end --> ends with
```

**Searching pattern allowing multiple characters**
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"pumpkin")

str_detect(string_vec,"Pumpkin")

str_detect(string_vec, "PUMPKIN")

str_detect(string_vec,"[Pp]umpkin") #allow either option
```

**Searching for multiple consecutive characters**
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z][a-zA-Z]") # starts with number then two letters
```

**Using . as wild card to capture all characters**
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # . is a special character that matches everything
```

**Escape character**
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[") # find [

str_detect(string_vec, "[\\[\\]]") # find either [ or ]
```

## **Factors ... **

```{r}
sex_vec = factor(c("male", "male", "female", "female"))

as.numeric(sex_vec)

```

**Relevel**
```{r}
sex_vec = fct_relevel(sex_vec, "male")
```







