---
title: "Tidy Data"
author: "ak5357"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(haven)
library(tidyverse)
```

This document will show how to tidy data.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names()
```


## Pivot Longer

This needs to go from wide to long format.

```{r}
pulse_tidy_df = 
  pulse_df |> 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m, #range of columns
    names_to = "visit", #original column names will go into this new column
    values_to = "bdi_score", #values will go to this new column
    names_prefix = "bdi_score_" #removes prefix from old column names
  ) |> 
  mutate(
    visit = replace(visit, visit == "bl", "00m") #replacing baseline visit with 00m
  ) |> 
  relocate(id, visit)


pulse_tidy_df
```


Would be better to do the upper two code chunks in one code chunk.

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = bdi_score_bl:bdi_score_12m, #range of columns
    names_to = "visit", #original column names will go into this new column
    values_to = "bdi_score", #values will go to this new column
    names_prefix = "bdi_score_" #removes prefix from old column names
  ) |> 
  mutate(
    visit = replace(visit, visit == "bl", "00m") #replacing baseline visit with 00m
  ) |> 
  relocate(id, visit) #puts id and visit columns first

pulse_df
```

Do one more exmaple.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv", na = c(".", "", " ")) |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = gd0_weight:gd18_weight,
    names_to = "gd_date",
    values_to = "weight"
  ) |> 
  mutate(
    gd_date = case_match(
      gd_date,
      "gd0_weight" ~ 0,
      "gd18_weight" ~ 18
    ))

litters_df
```


## Pivot wider

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control", "control"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 10, 4.2, 5)
  )

analysis_df
```

Pivot wider for human readability.
```{r}
analysis_df |> 
  pivot_wider(
    names_from = time,
    values_from = mean
  )
```


## Bind tables

This is where we'll learn to do knitr::kable()

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_ring, two_towers, return_king) |> 
  janitor::clean_names() |> 
  pivot_longer(
    cols = female:male,
    names_to = "sex",
    values_to = "words"
  ) |> 
  relocate(movie) |> 
  mutate(race = str_to_lower(race))
  
lotr_df
```


## Joining Tables

Import `litters` dataset.

```{r}
litters_df =
  read_csv("data/FAS_litters.csv", na = c(".", "", " ", "NA")) |> 
  janitor::clean_names() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  ) |> 
  separate(
    group,
    into = c("dose", "day_of_treatment"),
    sep = 3
  )

litters_df
```

Import `pups` next!
```{r}
pups_df = 
  read_csv("data/FAS_pups.csv", na = c(".", "", " ", "NA")) |> 
  janitor::clean_names() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )

pups_df
```

Join `litters` and `pups` dataframes together.

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df
```
